16s beta diversity

### 2021-03-08 scripts/48_16s_beta_diversity.Rmd
slowly but making progress

```{r}
# a function to do the combinations of the anosim test
pheno4 <- colnames(pheno)[2:5]

combi <- data_frame(
  g1 = c(rep('M',4), rep('P',4), rep('M',4)),
  g2 = c(rep('M',4), rep('P',4), rep('P',4)),
  label = rep(pheno4, 3)
)

do_anosim_each <-  function(g1_, g2_, label_){
    df = dat %>% 
      filter(center %in% c(g1_, g2_)) %>% 
      filter(get(label_) %in% c('yes','no')) 
    cts = df %>% 
      select(sampleid, all_of(keep)) %>% 
      column_to_rownames('sampleid')
    phe = df %>% 
      select(matches(label_)) %>% pull(get(label_)) 
    phedf = df %>% 
      select(sampleid, matches(label_))

@@

    # plot the pcoa
    dist_ <- vegdist(cts, method = 'bray')
    bc <- cmdscale(dist_, k = 2) 
    bc %>% 
      as.data.frame() %>% 
      rownames_to_column('sampleid') %>% 
      full_join(phedf, by = 'sampleid') %>% 
      ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
      labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}\npval:{ret$signif} statistic:{round(ret$statistic, 2)}')) +
      ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))

@@

>>>>

```{r}
# a function to do the combinations of the anosim test
pheno_ <- colnames(pheno)[2:5] 
n <- length(pheno_)

combi <- data_frame(
  g1 = c(rep('M',n), rep('P',n), rep('M',n)),
  g2 = c(rep('M',n), rep('P',n), rep('P',n)),
  label = rep(pheno_, 3)
)

do_anosim_each <-  function(g1_, g2_, label_){

    df = dat %>% 
      filter(center %in% c(g1_, g2_)) %>% 
      filter(get(label_) %in% c('yes','no')) 
    cts = df %>% 
      select(sampleid, all_of(keep)) %>% 
      column_to_rownames('sampleid')
    phe = df %>% 
      select(matches(label_)) %>% pull(get(label_)) 
    phedf = df %>% 
      select(sampleid, matches(label_))

@@

    # plot the pcoa
    dist_ <- vegdist(cts, method = 'bray')
    bc <- cmdscale(dist_, k = 2)
    bc %>%
      as.data.frame() %>%
      rownames_to_column('sampleid') %>%
      full_join(phedf, by = 'sampleid') %>%
      ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
      labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}\npval:{ret$signif} statistic:{round(ret$statistic, 2)}')) +
      theme_void() +
      ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))

@@

```{r}
# do the pcoa colored by the dominating genus 
g1_ = 'M'  
g2_ = 'P'
label_ = 'domgenus'
df = dat %>% 
  filter(center %in% c(g1_, g2_)) 
cts = df %>% 
  select(sampleid, all_of(keep)) %>% 
  column_to_rownames('sampleid')
phe = df %>% 
  select(matches(label_)) %>% pull(get(label_)) 
phedf = df %>% 
  select(sampleid, matches(label_))

# # plot the pcoa
dist_ <- vegdist(cts, method = 'bray')
bc <- cmdscale(dist_, k = 2)

# the version without the anosim stats
bc %>% 
  as.data.frame() %>% 
  rownames_to_column('sampleid') %>% 
  full_join(phedf, by = 'sampleid') %>% 
  ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
  labs(title = str_glue('Combined cohort colored by dominating genus')) +
  theme_void() +
  ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))
```
